---
title: "anthroDisturbance_Generator"
author: ""
date: "16 June 2022"
output:
  html_document:
    keep_md: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, results = "hold") ## TODO: use 'eval = TRUE'
```

# Overview

Provide an overview of what the module does / how to use the module.

Module documentation should be written so that others can use your module.
This is a template for module documentation, and should be changed to reflect your module.

## R Markdown

R Markdown syntax allows R code, outputs, and figures to be rendered in the documentation.

For help writing in R Markdown, see https://rmarkdown.rstudio.com/.

# Usage

```{r module_usage}
if(!require("Require")){
    install.packages("Require")
}
library("Require")
Require("googledrive")
Require("SpaDES.core")

# Pass your email for authentication (used for non-interactive calls)
googledrive::drive_auth(email = "tati.micheletti@gmail.com")
options(reproducible.useTerra = FALSE) # Workaround while reproducible is not yet fully functional with terra

# If you load the project, set the directory where your modules are located 
moduleDir <- dirname(getwd())
  
setPaths(modulePath = moduleDir,
         # cachePath = checkPath(file.path(getwd(), "cache"), 
         cachePath = checkPath(file.path(moduleDir, "potentialResourcesNT_DataPrep", "cache"), 
                               create = TRUE), 
         outputPath = checkPath(file.path(getwd(), "outputs"), 
                               create = TRUE),
         inputPath = checkPath(file.path(getwd(), "inputs"), 
                               create = TRUE),
         rasterPath = checkPath(file.path(getwd(), "temp_raster"), 
                               create = TRUE))

getPaths() # shows where the 4 relevant paths are



times <- list(start = 2011, end = 2041)

parameters <- list(
  anthroDisturbance_Generator = list(saveInitialDisturbances = FALSE)
  #.progress = list(type = "text", interval = 1), # for a progress bar
  ## If there are further modules, each can have its own set of parameters:
  #module1 = list(param1 = value1, param2 = value2),
  #module2 = list(param1 = value1, param2 = value2)
)
modules <- list("anthroDisturbance_Generator")
objects <- list()
inputs <- list()
outputs <- list()

disturbanceList <- simInitAndSpades(times = times, 
                                    params = parameters, 
                                    modules = modules,
                                    objects = objects)
```

# Parameters

Provide a summary of user-visible parameters.

```{r moduleParams, echo = FALSE, eval = TRUE}
df_params <- SpaDES.core::moduleParams("anthroDisturbance_Generator", "C:/Users/Tati/Documents/GitHub")
knitr::kable(df_params)
```

# Events

Describe what happens for each event type.

## Plotting

Write what is plotted.

## Saving

Write what is saved.

# Data dependencies

## Input data

How to obtain input data, and a description of the data required by the module.
If `sourceURL` is specified, `downloadData("anthroDisturbance_Generator", "C:/Users/Tati/Documents/GitHub")` may be sufficient.

```{r moduleInputs, echo = FALSE, eval = TRUE}
df_inputs <- SpaDES.core::moduleInputs("anthroDisturbance_Generator", "C:/Users/Tati/Documents/GitHub")
knitr::kable(df_inputs)
```

## Output data

Description of the module outputs.

```{r moduleOutputs, echo = FALSE, eval = TRUE}
df_outputs <- SpaDES.core::moduleOutputs("anthroDisturbance_Generator", "C:/Users/Tati/Documents/GitHub")
knitr::kable(df_outputs)
```

# Links to other modules

Describe any anticipated linkages to other modules.
